#Peer Assessment 1

Loading and transforming the data

```{r, echo = TRUE, highlight=TRUE}
act <- read.csv("activity.csv")
act$date<-as.factor(act$date)
head(act)
```

Histogram of the total number of steps per day:

```{r plot, echo=TRUE,highlight=TRUE}
new_act<-act[order(act$date),]
b<-split(new_act,new_act$date)
res<-sapply(b,function(elt) sum(elt[,1], na.rm=TRUE))
hist(res,main = "Histogram of the total number of steps", xlab="Steps", breaks=15)
```

Mean and median of the total number of steps per day:

```{r,echo=TRUE,highlight=TRUE}
x<-unique(new_act$date)
means<-sapply(b,function(elt) mean(elt[,1], na.rm=TRUE))
medians<-sapply(b,function(elt) median(elt[,1], na.rm=TRUE))
result1<-data.frame(x,means,medians)
names(result1) <- c("Date", "Mean of steps", "Median of steps")
print(result1)
```




Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r,echo=TRUE,highlight=TRUE}
new_act<-act[order(act$interval),]
b<-split(new_act,new_act$interval)
res<-sapply(b,function(elt) sum(elt[,1], na.rm=TRUE))
x<-unique(new_act$interval)
result<-data.frame(x,res)
names(result)<-c("Interval", "Steps_total")
with(result, plot(Interval, Steps_total, type = "l"))
```

Interval containig the maximum number of steps:

```{r,echo=TRUE,highlight=TRUE}
result[(result$Steps==max(result$Steps_total, na.rm = TRUE)),]
```

In this dataset there are 2304 rows with missing values:

```{r,echo=TRUE,highlight=TRUE}
sum(is.na(act$steps))
```


Filling in all of the missing values in the dataset (mean for this 5-minute interval).

```{r,echo=TRUE,highlight=TRUE}
tmp_act<-act
m <- merge(tmp_act, result, by.x = "interval", by.y = "Interval")
t<-m$steps
bad<-is.na(t)
t2<-m$Steps_total
t2<-t2[!bad]
t[is.na(t)]<-t2
m$steps<-t
summary(m)
```



Histogram and table with means and medians for new dataset:

```{r, echo=TRUE,highlight=TRUE}
new_act<-m[order(m$date),]
b<-split(new_act,new_act$date)
res<-sapply(b,function(elt) sum(elt[,2], na.rm=TRUE))
hist(res,main = "Histogram of the total number of steps", xlab="Steps", breaks=15)

x<-unique(new_act$date)
means<-sapply(b,function(elt) mean(elt[,2], na.rm=TRUE))
medians<-sapply(b,function(elt) median(elt[,2], na.rm=TRUE))
result1<-data.frame(x,means,medians)
names(result1) <- c("Date", "Mean of steps", "Median of steps")
print(result1)
```




As you can see histogram differs from the estimates in the 1st part. Mean's and median's values are practically the same. 
It's because of the fact that the total number of daily steps increase, but mean and median for each 5-minute interval stay the same.

We need to add column "weekday" to new dataset and make the plot with library lattice:

```{r, echo=TRUE,highlight=TRUE}
d<-as.POSIXlt(m$date)
m$weekday<-rep(0,nrow(m))
m$weekday[d$wday==0 | d$wday==6]<-"weekend"
m$weekday[m$weekday != "weekend"]<-"weekday"
m$weekday<-as.factor(m$weekday)
library(lattice)
m_wday<-m[(m$weekday=="weekday"),]
m_wend<-m[(m$weekday=="weekend"),]
b<-split(m_wday,m_wday$interval)
c<-split(m_wend,m_wend$interval)
res_wday<-sapply(b,function(elt) mean(elt[,2]))
res_wend<-sapply(c,function(elt) mean(elt[,2]))
x<-unique(m_wday$interval)
y<-unique(m_wend$interval)
wday<-data.frame(x,res_wday,rep("weekday",length(x)))
wend<-data.frame(y,res_wend,rep("weekend",length(y)))
names(wday)<-c("interval","steps", "weekday")
names(wend)<-c("interval","steps", "weekday")
new_act<-rbind(wday,wend)
xyplot(steps~interval|weekday,new_act, type="l")
```

